---
title: "exploring diagrammeR"
author: "Derek"
date: "April 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(DiagrammeR)

library(magrittr)
library(dplyr)


knitr::opts_chunk$set(echo = TRUE)
```


## tooling around

So eventually I want to take some tabular data, and make a workflow diagram out of that. DiagrammeR has GraphViz integration, which is neat, but I'm interested in scripting / programmatic creation of visualizations. 



From the documentation, it seems that in making a graph structure from tabular data, there needs to be a 1) Node Data Frame, and 2) an Edge Data Frame (NDF and EDF).

Let's mock up some data so I can play around with it.

```{r}
eat <- tibble::tribble(
         ~ID,              ~process, ~nextstep,
          1L,           "get bread",        2L,
          2L, "open jars of spread",        3L,
          3L,     "spread on bread",        4L,
          4L,                 "???",        5L,
          5L,              "PROFIT",        NA
         )

```


From this example I have a simple process, and a column denoting the next step.

Let's make an NDF and EDF from this. 

```{r, eval = FALSE}
create_nodes(nodes = eat$process)
# Error in create_nodes(nodes = eat$process) : 
#   could not find function "create_nodes"
```
 
 *Great, the documentation doesn't match*
 
Also, what's kind of weird is that the GitHub readme has another way of thinking about creating graphs, using a more pipe-oriented approach: https://github.com/rich-iannone/DiagrammeR

Looks like `create_node_df` is my closest bet -- let's read the help content

```{r}
eat_node <- create_node_df(n = length(eat$process), label = eat$process)

eat_edge <- eat %>% filter(!(is.na(ID)), !(is.na(nextstep))) %>% 
  with(., create_edge_df(from = .$ID, to = .$nextstep, rel = "next step"))
```

Blend them together

```{r}
create_graph(eat_node, eat_edge) %>% render_graph()
```


## Fancier graph

```{r}
eat <- tibble::tribble(
         ~ID, ~process,                         ~step, ~nextstep,
          1L,  "bread",                   "get bread",        2L,
          2L,  "bread",              "place 2 slices",        3L,
          3L, "spread",             "get two spreads",        4L,
          4L, "spread",         "open jars of spread",        5L,
          5L, "spread",     "spread one on one bread",        7L,
          6L, "spread", "spread other on other bread",        7L,
          7L,   "prep",                 "close bread",        8L,
          8L, "finish",                         "???",       9L,
         9L, "finish",                      "PROFIT",        NA
         )

```


```{r}
eat_node <- create_node_df(n = length(eat$step), label = eat$step, type = eat$process)

eat_edge <- eat %>% filter(!(is.na(ID)), !(is.na(nextstep))) %>% 
  with(., create_edge_df(from = .$ID, to = .$nextstep, rel = "next step"))
```

Blend them together

```{r}
create_graph(eat_node, eat_edge) %>% render_graph()
```

Ah, this is going to be a little tricky to get a graph to split and come back together (e.g. parallel steps).


I'm thinking that I'm going to have to do something to capture how processes can split, capture this in a table, and then easily, and then somehow build edges from this. 



**Noodling around some more I found a different version of the docs**: http://visualizers.co/diagrammer/reference/index.html


## What graphing packages are available?

- DiagrammeR (from RStudio)
  - last development done in ~ 6 months ago
- igraph (https://igraph.org/r/)
- networkD3

